Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PS lipids

### 2018-11-20 Manuscript/manuscriptPS.tex
Update text

\begin{figure*}[]
  \centering
  \includegraphics[width=18.0cm]{../Figs/figS8.png}
  \caption{\label{HGstructuresPS}
    Overlayed snapshots from glycerol backbone and headgroup region from 
    different simulations of PS lipids.
  }
\end{figure*}

@@

The headgroup and glycerol backbone of PS lipids show wide variety between different simulation
models in the order parameters and structures (Figs. \ref{HGorderParametersPS} and \ref{HGstructuresPS}),
as previously observed also for PC lipids \cite{botan15}.
The models perform generally less well for PS lipids than for PC lipids in the previous study
(Figs. 2 and 4 in Ref.~\cite{botan15} vs. Figs.~\ref{HGorderParametersPS} and \ref{comparisonTablePS}).
Therefore, interpretation of structural differences between PC and PS headgroups
from simulations is not straightforward.

The best performing models,
Slipids, CHARMM36 and CHARMM36ua, reproduce the larger forking of the $\alpha$-carbon 
and the Slipids model reproduces also the lower of the $\beta$-carbon order parameter
when comparing the PS results to PC (Fig. 2 in Ref.~\citenum{botan15} vs. Fig. \ref{HGorderParametersPS}).
Interestingly, the dihedral angle distributions of C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ show
a single narrow maximum close to 120$^{\circ}$ in the best three models, while other models give several maxima
in different locations (Fig. \ref{dihedralsHG}). The restricted motion is also
visible in the sampled conformations (Fig.~\ref{HGstructuresPS}) and might
manifest the increased rigidity anticipated from the early experimental studies \cite{browning80,buldt81}.
\todo{We should consider making the conformation figures from CHARMM36 and Slipids POPC, and compare these
  to the results from POPS.}

\todo{Maybe we should move the glycerol backbone results to SI and focus here on the headgroup,
and especially to the difference between PC and PS.}
The glycerol backbone order parameters of C$_2$ and C$_3$ from Slipids simulations
differ significantly from the other simulation results and experiments (Fig. \ref{HGorderParametersPS}),
as observed previously also for PC lipids \cite{botan15}.
The origin of this difference is more difficult to track without more elaborate analysis,
because different models show very complicated patterns of distinct structures
in the glycerol backbone region (Figs.~\ref{HGstructuresPS} and \ref{dihedralsHG}).

>>>>

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/structures.eps}
  \caption{\label{HGstructuresPSandPC}
    Overlayed snapshots from glycerol backbone and headgroup region from 
    different simulations of PS lipids.
  }
\end{figure}

@@

The headgroup and glycerol backbone of PS lipids show wide variety between different simulation
models in the order parameters and structures (Figs. \ref{HGorderParametersPS} and \ref{HGandGLYstructuresPS}),
as previously observed also for PC lipids \cite{botan15}.
The models perform generally less well for PS lipids than for PC lipids in the previous study
(Figs. 2 and 4 in Ref.~\cite{botan15} vs. Figs.~\ref{HGorderParametersPS} and \ref{comparisonTablePS}).
Therefore, interpretation of structural differences between PC and PS headgroups
from simulations is not straightforward.

The best performing models,
Slipids, CHARMM36 and CHARMM36ua, reproduce the larger forking of the $\alpha$-carbon 
and the Slipids model reproduces also the lower of the $\beta$-carbon order parameter
when comparing the PS results to PC (Fig. 2 in Ref.~\citenum{botan15} vs. Fig. \ref{HGorderParametersPS}).
Interestingly, the dihedral angle distributions of C$_\alpha$-C$_\beta$-C$_\gamma$-O$_\gamma$ show
a single narrow maximum close to 120$^{\circ}$ in the best three models, while other models give several maxima
in different locations (Fig. \ref{dihedralsHG}). The restricted motion is also
visible in the sampled conformations (Figs.~\ref{HGstructuresPSandPC} a) and \ref{HGandGLYstructuresPS})
suggesting that the rotation of carboxyl group is limited in the serine headgroup.
In addition, the comparison of N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral between PC and PS headgroups
from CHARMM36 simulations, which are in best agreement with experiments for both lipids,
shows a more asymmetric and restricted
rotations around this bond (Figs.~\ref{HGstructuresPSandPC} b) and \ref{dihedralsHGpc}).
These results might manifest the increased rigidity anticipated from the early experimental studies \cite{browning80,buldt81}.
Also the dihedral distributions and sampled conformations of glycerol backbone region
from different models are shown in the supplementary information
(Figs. \ref{dihedralsGLY}, \ref{HGandGLYstructuresPS}, \ref{dihedralsGLYpc} and \ref{HGandGLYstructuresPSPC}),
but further analysis is beyond the scope of this work which the PS headgroup.

### 2018-11-20 Manuscript/manuscriptPSsuppl.tex
Update text

@@

>>>>

The glycerol backbone order parameters of C$_2$ and C$_3$ from Slipids simulations
differ significantly from the other simulation results and experiments (Fig. \ref{HGorderParametersPS}),
as observed previously also for PC lipids \cite{botan15}.
The origin of this difference is more difficult to track without more elaborate analysis,
because different models show very complicated patterns of distinct structures
in the glycerol backbone region (Figs.~\ref{HGstructuresPS} and \ref{dihedralsHG}).

@@

\begin{figure*}[]
  \centering
  \includegraphics[width=18.0cm]{../Figs/figS8.png}
  \caption{\label{HGandGLYstructuresPS}
    Overlayed snapshots from glycerol backbone and headgroup region from 
    different simulations of PS lipids.
  }
\end{figure*}

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/figS7PC.png}
  \caption{\label{dihedralsHGpc}
    Dihedral angle distributions of headgroup region of PS lipids from different simulation models.
  }
\end{figure*}

\begin{figure*}[]
  \centering
  \includegraphics[width=15.0cm]{../Figs/figS6PC.png}
  \caption{\label{dihedralsGLYpc}
    Dihedral angle distributions of glycerol backbone region of PS lipids from different simulation models.
  }
\end{figure*}

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/figS8_POPC.png}
  \caption{\label{HGandGLYstructuresPSPC}
    Overlayed snapshots from glycerol backbone and headgroup region from 
    different simulations of PS lipids.
  }
\end{figure}
